---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

Load needed packages

```{r}
libs <- c('dplyr', 'moments', 'ggplot2', 'latticeExtra', 'gridExtra', 'MASS', 
          'colorspace', 'plyr', 'Hmisc', 'scales')
invisible(lapply(libs, require, character.only = T))
```
Display Environment
```{r}
sessionInfo()

```

Unzip and Load the data
```{r}
setwd("C://Users//KathyWarner.KATHYSDESKTOP//Documents//GitHub//RepData_PeerAssessment1")

fileURL<-"./activity.zip"

unzip(zipfile="./activity.zip")

activity <- read.csv("./activity.csv")

```

Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day

```{r histogram}
Totalsteps<-with(activity,(tapply(steps,date,sum, na.rm=TRUE)))

hist(Totalsteps, xlab="Total Number of Steps Taken Each Day", main="Total Steps")


mean(Totalsteps, na.rm=TRUE)
median(Totalsteps, na.rm=TRUE)
```

Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

```{r}

averages <- aggregate(x=list(steps=activity$steps), by=list(interval=activity$interval),
                      FUN=mean, na.rm=TRUE)
require(ggplot2)
ggplot(data=averages, aes(x=interval, y=steps)) +
        geom_line() +
        xlab("5-minute Intervals") +
        ylab("Average Number of Steps Taken")

```

Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
topAvg<-averages[which.max(averages$steps),]
topAvg
```

Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)

```{r}
NArows<-apply(activity,1, function(x) is.na(x))

NArows1<-sum(NArows)

```
The total number of rows with NAs is `r NArows1`.


The strategy for filling in all of the missing values in the dataset is to use the mean for that 5-minute interval. A new dataset is created that is equal to the original dataset but with the missing data filled in.

```{r Merge}
averages<-rename(averages, c("steps"="isteps"))
newData<-merge(activity, averages, by="interval")

newData$steps <- ifelse(is.na(newData$steps), newData$isteps, newData$steps)
newData<-newData[,1:3]        

```

Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. Do these values differ from the estimates from the first part of the assignment? 

What is the impact of imputing missing data on the estimates of the total daily number of steps?

```{r}
Totalsteps1<-with(newData,(tapply(steps,date,sum, na.rm=TRUE)))

hist(Totalsteps1, xlab="Total Number of Steps Taken Each Day", main="Total Steps")

mean(Totalsteps1, na.rm=TRUE)
median(Totalsteps1, na.rm=TRUE)
```
The impact of imputing missing data on the estimates of the total daily number of steps is an increase in the mean and the median.  Now the mean and median are equal which results in a more symmetrical histogram. 


Create a new factor variable in the dataset with two levels -- "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.
```{r}
library(lubridate)
newData$date <- as.Date(newData$date)

newData$week<-ifelse(wday(newData$date, label=TRUE)  %in% c("Mon", "Tues", "Wed", "Thurs", "Fri"), 'Weekday', 'Weekend')
newData$steps <- ifelse(is.na(newData$steps), newData$isteps, newData$steps)
newData$week<-as.factor(newData$week)
```

Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). The plot should look something like the following, which was created using simulated data:
```{r }
newData1 <- aggregate(steps ~ interval + week, data=newData, mean)

class(newData1)

g <- ggplot(newData1, aes(x =interval, y = steps))
g + geom_line() + facet_wrap(~week) + 
    labs(x = "5-Minute Interval") + 
    labs(y = "Average Number of Steps") 
```